---
title: "Effects of urban green infrastructure on biodiversity: birds (?)"
author: "Bianca Neves & Irina Lerner"
date: "19/06/24"
output:
  html_document:
    toc: yes
    toc_float: yes
    theme: paper
  pdf_document:
    toc: yes
---

```{r setup, include=FALSE, echo=FALSE}
knitr::opts_chunk$set(echo = TRUE)
#source("functions2.R")
library(ggplot2)
library(PerformanceAnalytics) # for correlation matrix
library(tidyr) # for organized code
library(sf, sp) # for reading shp and ploting
library(spdep)
library(glmnet)
# customs
custom_theme <- theme_minimal() +
                theme(text = element_text(size = 12),
                      plot.title = element_blank(),
                      panel.grid.major = element_blank(),
                      panel.grid.minor = element_blank())
```


# Data

We begin with a $1km^2$ grid over the urban area of the city of São Paulo

```{r data, include=FALSE}
# sum of species
shp <- st_read("../data/1km/richness_effort_v6_total_1km_urb_f1km_30min.shp")
cdata <- shp
# Normalize by log 
cdata$re <- cdata$richness/log(cdata$effort)

AIC(lm(richness ~ 1, cdata))
AIC(lm(re ~ 1, cdata))

# presence-absence table
sp <- st_read("../data/1km/ebd_total_v5_1km_urb_f1km_30min.shp")
# Create a table of unique species and grid IDs
unique_species <- unique(sp$scntfc_)
unique_ids <- unique(sp$id)

# Initialize the presence-absence matrix with zeros
presence_absence <- matrix(0, nrow = length(unique_ids), ncol = length(unique_species))
colnames(presence_absence) <- unique_species
rownames(presence_absence) <- unique_ids

# Fill the presence-absence matrix
for (i in 1:nrow(sp)) {
  grid_id <- sp$id[i]
  species <- sp$scntfc_[i]
  presence_absence[as.character(grid_id), as.character(species)] <- 1
}

# Convert the matrix to a data frame for better readability
PA <- as.data.frame(presence_absence)
```

Let's see it as a ternary plot

```{r}
library(ggtern)
library(viridis)


ggtern(data = cdata, aes(x = SVI, y = SGVI, z = SUVI, color = re)) +
  geom_point(size = 3) +  # Adjust point size if needed
  scale_color_viridis_c() +  # Color gradient
  theme_bw() +  # Use a black and white theme
  labs(title = "oi", x = "SVI", y = "SGVI", z = "SUVI", color = "Richness")

```


We can see that normalizing by re is adequate

```{r}
AIC(lm(richness ~ 1, cdata))
AIC(lm(re ~ 1, cdata))
```

So we proceed with richness normalized by effort.

There is high correlations among variables

```{r}
cor_mat <- as.data.frame(cdata)[c("richness", "effort", "re", "SVI", "SGVI", "SUVI", "SEVI", "prop_veg", 
                                  "Bheight_me", "Vheight_me", "Bvol", "Vvol", "Wdistance")]
suppressWarnings(chart.Correlation(cor_mat, histogram=TRUE, pch=16, method = "spearman"))
```

Testing for spatial correlation

```{r}
coords <- st_centroid(cdata)
nb <- poly2nb(cdata)
lw <- nb2listw(nb, style="W", zero.policy=TRUE)
moran.mc(cdata$re, lw, nsim = 999)
```
We see that there is spatial correlation

Lets check the spatial distribuition of re

```{r}
ggplot(data = cdata) +
  geom_sf(aes(fill = re), color = NA) +
  scale_fill_viridis_c(direction = -1) +
  custom_theme
```

# Linear variable selection

We use lasso and ridge regression, a method that penalizes variables for multicolinearity and selects significant variables.

```{r}
variables <- c("SVI", "SGVI", "SUVI", "SEVI", "prop_veg", "Bheight_me", "Vheight_me", "Bvol", "Wdistance")

cdata[is.na(cdata)] <- 0

X <- as.matrix(as.data.frame(cdata)[, variables])
re <- cdata$re

lasso_model <- cv.glmnet(X, re, alpha = 1)
ridge_model <- cv.glmnet(X, re, alpha = 0)

lasso_coefs <- coef(lasso_model, s = "lambda.min")
ridge_coefs <- coef(ridge_model, s = "lambda.min")

# Prepare data for plotting
variable_names <- rownames(lasso_coefs)[-1]  # Exclude intercept
lasso_effect_sizes <- lasso_coefs[-1, 1]  # Exclude intercept
ridge_effect_sizes <- ridge_coefs[-1, 1]  # Exclude intercept

# Create a data frame
effect_sizes_df <- data.frame(
  Variable = variable_names,
  Lasso = lasso_effect_sizes,
  Ridge = ridge_effect_sizes
)

# Reshape data for ggplot2
effect_sizes_long <- gather(effect_sizes_df, key = "Model", value = "EffectSize", -Variable)

# Create the plot
ggplot(effect_sizes_long, aes(x = EffectSize, y = Variable, color = Model)) +
  geom_point(size = 3) +
  geom_vline(xintercept = 0, linetype = "dashed", color = "grey") +
  labs(x = "Effect Size", y = "") +
  theme_minimal() +
  theme(axis.text.y = element_text(size = 10))

```
With this, we show that SVI and SEVI are a good method for analysing bird diverisity in the 1km range

```{r}
library(ggeffects)
full_model <- lm(re ~ SVI + SEVI, data = cdata)

summary(full_model)


# Get the summary of the model
model_summary <- summary(full_model)
coefficients <- model_summary$coefficients[, "Estimate"]
standard_errors <- model_summary$coefficients[, "Std. Error"]

# Combine them into a data frame
effects_df <- data.frame(
  Variable = names(coefficients),
  EffectSize = coefficients,
  StdError = standard_errors
)

# Create the plot
ggplot(effects_df[-1,], aes(x = EffectSize, y = Variable)) +
  geom_point(size = 3) +
  geom_vline(xintercept = 0, colour="red", linetype = "dashed") +
  geom_errorbarh(aes(xmin = EffectSize - StdError, xmax = EffectSize + StdError), height = 0.2) +
  labs(x = "Effect Size", y = "") +
  theme_bw()
```

```{r}
library(spaMM)

centroids <- st_centroid(cdata)
cdata$centx <- st_coordinates(centroids)[, 1]
cdata$centy <- st_coordinates(centroids)[, 2]

ftmod <- fitme(re ~ SVI + SEVI + Matern(1 | centx + centy), data = cdata, method = "ML")
AIC.HLfit(ftmod)
```
We see a small improvement in relation to the linear model without space.


```{r}
library(mgcv)

gammod <- gam(re ~ s(SVI, SEVI), data = cdata, method = "ML")
summary(gammod)
AIC(gammod, full_model)
vis.gam(gammod, view = c("SVI", "SEVI"),
    theta = 35, n.grid = 50, lwd = 0.4)
```

This analysis indicates non linearity.

# Random forest

```{r}
library(randomForest)

data_rf <- data.frame(re = cdata$re,
                      PV = cdata$prop_veg,
                      SGVI = cdata$SGVI,
                      SVI = cdata$SVI,
                      SUVI = cdata$SUVI,
                      SEVI = cdata$SEVI,
                      BV = cdata$Bvol,
                      VV = cdata$Vvol,
                      BH = cdata$Bheight_me,
                      VH = cdata$Vheight_me,
                      W = cdata$Wdistance,
                      coord_x = cdata$centx,
                      coord_y = cdata$centy)

# Fit Random Forest model
rf_model <- randomForest(re ~ ., data = data_rf, importance = TRUE)

# Summarize the model
print(rf_model)
# Check variable importance
importance <- rf_model$importance
print(importance)

# Plot variable importance
varImpPlot(rf_model)

# Create a plot for variable importance
importance_df <- data.frame(
  Variable = rownames(importance),
  Importance = importance[,1]
)

ggplot(importance_df, aes(x = reorder(Variable, Importance), y = Importance)) +
  geom_bar(stat = 'identity') +
  coord_flip() +
  labs(title = 'Variable Importance', x = 'Variables', y = 'Importance') +
  custom_theme
```

```{r}
library(pdp)

# Partial Dependence Plot for a single feature
pdp_pv <- partial(rf_model, pred.var = "BV", plot = TRUE, rug = TRUE, train = data_rf)
print(pdp_pv)

# For multiple features
pdp_sgvi <- partial(rf_model, pred.var = "PV", plot = TRUE, rug = TRUE, train = data_rf)
print(pdp_sgvi)

# For multiple features
pdp_svi <- partial(rf_model, pred.var = "SVI", plot = TRUE, rug = TRUE, train = data_rf)
print(pdp_svi)

# For multiple features
pdp_sevi <- partial(rf_model, pred.var = "SEVI", plot = TRUE, rug = TRUE, train = data_rf)
print(pdp_sevi)
```

```{r}
library(vegan)

PAid <- PA
PAid$id <- as.numeric(rownames(PA))

PA_df <- PAid[order(PAid$id),]


env.data <-data.frame(id = cdata$id,
                      PV = cdata$prop_veg,
                      SGVI = cdata$SGVI,
                      SVI = cdata$SVI,
                      SUVI = cdata$SUVI,
                      SEVI = cdata$SEVI,
                      BV = cdata$Bvol,
                      VV = cdata$Vvol,
                      BH = cdata$Bheight_me,
                      VH = cdata$Vheight_me,
                      W = cdata$Wdistance)

env_data <- env.data[order(env.data$id),]

rda_model <- rda(PA_df ~ ., data = env_data[,-1])

# Plot the RDA results
plot(rda_model, scaling = 2)
```
